Spin canonical divisors of log stacky curves

### 2015-06-29 stacky_write_up/spin_cring.tex
address o'dorney in intro, other intro edits

\section{Introduction}
Let $\Gamma$ be a {\bf Fuchsian} group, \todo{Peter: comma or colon here?} i.e. a discrete subgroup
of $PSL_2(\BR)$ acting on the upper half plane $\bh$ by fractional
linear transformations. We consider the {\bf ring of modular forms}
$M(\Gamma) = \bigoplus_k M_k(\Gamma)$. In this paper, we sharply bound the degree of the generators and
relations for $M(\Gamma)$, for general $\Gamma$ \todo{Peter: Why do you say general gamma?  I think you should either add "general fuchsian groups gamma" or just take out that clause, since you just defined what gamma is.}  \todo{Peter: also, its not clear to me that this should be said right here.  I think we may want to say a bit more about what's already known instead of going directly to what we have done}.  In the special that $\Gamma$ has no odd weight modular forms, Voight and Zureick-Brown
\cite[Chapters 7-9]{vzb:stacky} give similar bounds, and we extend their work to all Fuchsian groups $\Gamma$. In order to find generators and relations for $M(\Gamma)$, we translate the seemingly analytic question of understanding the ring of modular forms into the algebraic
category, using a generalization of the GAGA principle. By
\cite[Proposition 6.1.5]{vzb:stacky}, there is an equivalence of
categories between orbifold curves and log stacky curves over $\BC$
. Hence, for the remainder of the paper, we will work in the
algebraic category.

@@

\todo{Peter: I think we need something about odd weight modular forms before going right into stacks.}  
Following Voight and Zureick-Brown \cite{vzb:stacky}, we generalize Petri's theorem in many \todo{Peter: not clear to me that ``many'' is a good word to use here.  We mostly generalize in one direction}
directions. 
If $X$ is a curve, \todo{why do you start with a curve and then move to a stacky curve? shouldn't it be the opposite order?} then we say $\sx$ is a {\bf stacky curve} with {\bf coarse space} $X$ and {\bf stacky points} $P_1,\ldots, P_r$ with {\bf stabilizer orders} $e_1,\ldots, e_r \in \BZ_{\geq 2}$ if 
$$\di \sx = \di X \oplus \left(\bigoplus_{i
= 1}^r \left\langle \frac{1}{e_i}P_i \right\rangle \right) \subset \BQ \otimes \di X.$$
As is done in Voight and Zureick-Brown \cite{vzb:stacky} one may view a stacky curve as a stack, but for the purposes of this paper, it suffices to view $\sx$ as a scheme whose associated divisor group allows a finite number of ``fractional points,'' i.e., stacky points. A {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$ where $\Delta \in \di X$ is an effective divisor and $2\halfcan = K_X + \Delta + \sum_{i=1}^{r}\subhalf{e_i}P_i.$
The central object of study in this paper is the {\bf log spin
canonical ring} of $(\sx,\Delta,L)$ is
\begin{align*}
	R(\sx,\Delta,L) = \bigoplus_{d \geq 0} H^0(\sx,\lfloor d L \rfloor)
\end{align*}

\noindent
The main theorem of this paper is the following bound on the degree
of generators and relations of arbitrary log spin canonical rings:

@@

@@

\begin{rem}
\label{rem:explicit-generators}
The proofs of the main theorems in the low genus cases not only
obtain bounds on the degrees of the generators and relations, but
also yield explicit systems of generators and the initial ideal of
relations. Furthermore, our proof of the high genus case also
provides an inductive procedure for determining the generators and
initial ideal of relations for the half-canonical divisor $\halfcan$ can be written as $\halfcan = \halfcan' + \Lambda$ with
$(\halfcan', \Delta, \sx)$ when $\Lambda \in \di \sx$ effective. In fact, such a system of
generators and relations are known for $\halfcan'$ in many interesting low genus cases,
as detailed in Neves ~\cite[ Section III.4]{neves:halfcan}.
\end{rem}

@@

\todo{incorporate some chunk on O'Dorney somewhere in intro, below is previously written stuff}
In \cite{dorney:canonical}, O'Dorney uses best lower approximations to provide a minimal presentation for the canonical ring of a divisor $D = \frac{\alpha_1}{\beta_1}P_1 + \frac{\alpha_2}{\beta_2}P_2$ of $\BP_k^1$ \todo{This comes up a lot, but instead of saying $D$ is a divisor of $X$, I'd prefer to say $D \in \di X$ to be more precise. Of course, this is minor, and it's fine to keep it this way.} with at most 2 points in its support, \todo{It might be clearer to say $\# \Supp(D) \leq 2.$} to show it is generated in degree $e=\max{\beta_1,\beta_2}$ with relations generated in degrees at most $2e$. \todo{Can you break the previous sentence into two sentences?} He \todo{Probably better not to say he, maybe give a theorem number where he does this?} further extends the result to more points, but only bounds the degree of minimal generation by a value on the order the least common multiple of the denominators of fractional points. \todo{This sentence seems very imprecise to me, and I wouldn't understand it if I hadn't read Evan's paper. Can you state this more precisely, or possibly just remove it.}

In Lemma ~\ref{lem:sat-1} we provide an inductive method to strengthen
O'Dorney's result \cite[Theorem 8]{dorney:canonical} on $\BP^1_k$ to divisors $D = \sum_{i=1}^n \frac{\alpha_i
}{\beta_i} P_i$ for points $P_1, \ldots, P_n$ on a curve $X$ such that $H^
0(X, \lfloor D \rfloor) > 0$. Our result implies that $R_D$ is generated in
degrees at most $e = \max(\beta_i)$ with relations generated in
degrees at most $2e$. \todo{Are you sure you're not assuming $\frac{\alpha}{\beta}$ is effective? If not, doesn't what you say contradict Evan's example showing that his bounds of about lcm are asymptotically sharp?} Furthermore, given additional constraints on $H^0(
X, \lfloor dD \rfloor)$ for various values of $d$, we can apply the
same Riemann-Roch \todo{same as what? Maybe just say we can generalize his results to arbitrary genus?} arguments in arbitrary genus. 
To do this, we extend an inductive result \todo{Is ``inductive result'' a phrase? Can we just remove the word ``inductive'' every time it appears?} by Voight and
Zureick-Brown \cite[Theorem 8.3.1]{vzb:stacky} on stacky canonical rings to arbitrary $\BQ$-
divisors. \todo{Again, I don't think it can be arbitrary, and what are the $\BQ$ divisors on?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Background %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Somewhere in the paper, we should probably give an example showing the generators and relations are sharply bounded at degree e, 2e. Robin: Aren't the base cases suitable examples? Aaron: Those don't show it for general e, do they? I don't think we really need an example, just a remark somewhere that one cannot really hope to do better than e, 2e somewhere.}

@@

\todo{do this part}
Part on Kodaira-Spencer \cite{kodaira:complex-manifolds}.
Use \cite[Lemma 10.2.1]{vzb:stacky} to obtain the desired translation to algebraic 
geometry.

>>>>

\section{Introduction}
Let $\Gamma$ be a {\bf Fuchsian} group, i.e. a discrete subgroup
of $PSL_2(\BR)$ acting on the upper half plane $\bh$ by fractional
linear transformations. We consider the {\bf ring of modular forms}
$M(\Gamma) = \bigoplus_k M_k(\Gamma)$. In this paper, we sharply bound the degree of the generators and
relations for $M(\Gamma)$, for general $\Gamma$ \todo{Peter: Why do you say general gamma?  I think you should either add "general fuchsian groups gamma" or just take out that clause, since you just defined what gamma is.}  \todo{Peter: also, its not clear to me that this should be said right here.  I think we may want to say a bit more about what's already known instead of going directly to what we have done}.  In the special that $\Gamma$ has no odd weight modular forms, Voight and Zureick-Brown
\cite[Chapters 7-9]{vzb:stacky} give similar bounds, and we extend their work to all Fuchsian groups $\Gamma$. In order to find generators and relations for $M(\Gamma)$, we translate the seemingly analytic question of understanding the ring of modular forms into the algebraic
category, using a generalization of the GAGA principle. By
\cite[Proposition 6.1.5]{vzb:stacky}, there is an equivalence of
categories between orbifold curves and log stacky curves over $\BC$
. Hence, for the remainder of the paper, we will work in the
algebraic category.

@@

\todo{Peter: I think we need something about odd weight modular forms before going right into stacks.}  
Following Voight and Zureick-Brown \cite{vzb:stacky}, we generalize Petri's theorem in many \todo{Peter: not clear to me that ``many'' is a good word to use here.  We mostly generalize in one direction}
directions. 
If $X$ is a curve, \todo{why do you start with a curve and then move to a stacky curve? shouldn't it be the opposite order?} then we say $\sx$ is a {\bf stacky curve} with {\bf coarse space} $X$ and {\bf stacky points} $P_1,\ldots, P_r$ with {\bf stabilizer orders} $e_1,\ldots, e_r \in \BZ_{\geq 2}$ if 
$$\di \sx = \di X \oplus \left(\bigoplus_{i
= 1}^r \left\langle \frac{1}{e_i}P_i \right\rangle \right) \subset \BQ \otimes \di X.$$
As is done in Voight and Zureick-Brown \cite{vzb:stacky} one may view a stacky curve as a stack, but for the purposes of this paper, it suffices to view $\sx$ as a scheme whose associated divisor group allows a finite number of ``fractional points,'' i.e., stacky points. A {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$ where $\Delta \in \di X$ is an effective divisor and $2\halfcan = K_X + \Delta + \sum_{i=1}^{r}\subhalf{e_i}P_i.$
The central object of study in this paper is the {\bf log spin
canonical ring} of $(\sx, \Delta, \halfcan)$ is
\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(\sx,\lfloor d \halfcan \rfloor)
\end{align*}

\noindent
\todo{incorporate some chunk on O'Dorney somewhere in intro, below is previously written stuff}
In particular, we would like to bound the degrees in which the
graded ring and its relations are generated. The work by O'Dorney
shows that the canonical ring of any $\BQ$-divisor $D =
\frac{\alpha_1}{\beta_1}P_1 + \frac{\alpha_2}{\beta_2}P_2$ of
$\BP_k^1$ with $\# \Supp(D) \leq 2$ is generated in degrees up
to $e := \max(\beta_1, \beta_2)$ with relations generated in
degrees up to $2e$ \cite[Section 4]{dorney:canonical}. Furthermore,
O'Dorney gives a larger bound when the support of $D$ is allowed to
have arbitrarily many points in terms of least-common-multiples.
Voight and Zureick-Brown also bounded the generator and relation
degrees of log spin canonical rings of $(\sx, \Delta, \halfcan)$,
in the case that genus is positive and $\halfcan$ is effective,
by $3e$ and $6e$ for $e := \max(e_1, \ldots, e_r)$.

The goal of this paper is to provide a stricter bound on the
generator and relation degrees of log spin canonical rings
for all genera. Note that O'Dorney's bounds are asymptotically
strict for arbitrary $\BQ$-divisors on $\BP^1$, but restricting
to the log spin curve case allows for lower bounds.
The main theorem of this paper is the following improved bound on
the degree of generators and relations of log spin canonical rings:

%uses best lower approximations to provide a minimal presentation for the canonical ring of an arbitrary $\BQ$-divisor $D = \frac{\alpha_1}{\beta_1}P_1 + \frac{\alpha_2}{\beta_2}P_2$ of $\BP_k^1$ with $\# \Supp(D) \leq 2$ and shows that it is generated in degree $e := \max(\beta_1, \beta_2)$ with relations generated in degrees at most $2e$ \cite[Section 4]{dorney:canonical}. For arbitrarily many points, O'Dorney also provides much larger bounds \cite[Section 5]{dorney:canonical}. While
%
%In Lemma ~\ref{lem:sat-1} we provide an inductive method to strengthen
%O'Dorney's result \cite[Theorem 8]{dorney:canonical} on $\BP^1_k$ to divisors $D = \sum_{i = 1}^n \frac{\alpha_i
%}{\beta_i} P_i$ for points $P_1, \ldots, P_n$ on a curve $X$ such that $H^
%0(X, \lfloor D \rfloor) > 0$. Our result implies that $R_D$ is generated in
%degrees at most $e = \max(\beta_i)$ with relations generated in
%degrees at most $2e$. \todo{Are you sure you're not assuming $\frac{\alpha}{\beta}$ is effective? If not, doesn't what you say contradict Evan's example showing that his bounds of about lcm are asymptotically sharp?} Furthermore, given additional constraints on $H^0(
%X, \lfloor dD \rfloor)$ for various values of $d$, we can apply the
%same Riemann-Roch \todo{same as what? Maybe just say we can generalize his results to arbitrary genus?} arguments in arbitrary genus. 
%To do this, we extend a result by Voight and
%Zureick-Brown \cite[Theorem 8.3.1]{vzb:stacky} on stacky canonical rings to arbitrary $\BQ$-
%divisors. \todo{Again, I don't think it can be arbitrary, and what are the $\BQ$ divisors on?}

@@

\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(\sx, \lfloor d \halfcan \rfloor)
\end{align*}

@@

\begin{rem}
\label{rem:explicit-generators}
The proofs of the main theorems in the low genus cases not only
obtain bounds on the degrees of the generators and relations, but
also yield explicit systems of generators and the initial ideal of
relations. Furthermore, our proof of the high genus case also
provides an inductive procedure for determining the generators and
initial ideal of relations for the half-canonical divisor $\halfcan$
 can be written as $\halfcan = \halfcan' + \Lambda$ with $(\halfcan
', \Delta, \sx)$ when $\Lambda \in \di \sx$ effective. In fact,
such a system of generators and relations are known for $\halfcan'$
in many interesting low genus cases, as detailed in Neves
~\cite[Section III.4]{neves:halfcan}.
\end{rem}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Background %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\todo{Somewhere in the paper, we should probably give an example showing the generators and relations are sharply bounded at degree e, 2e. Robin: Aren't the base cases suitable examples? Aaron: Those don't show it for general e, do they? I don't think we really need an example, just a remark somewhere that one cannot really hope to do better than e, 2e somewhere.}

@@

\todo{do this part}
Part on Kodaira--Spencer \cite{kodaira:complex-manifolds}.
Use \cite[Lemma 10.2.1]{vzb:stacky} to obtain the desired translation to algebraic 
geometry.
